Title Goes Here \thanks{The research was supported by UBACyT20020170100428BA and the CLIMAX Project funded by Belmont Forum/ANR-15-JCL/-0002-01. Elio Campitelli was supported by a PhD grant from CONICET, Argentina.}

### 2021-04-08 analysis/paper/shceof.tex
Renderiza todo

\textcolor{red}{(La discusión de los próximos parráfos me resulta confusa. Deberías aclarar mejor por qué mostras cada cosa que mostras y cómo se relaciona con lo que mostraste antes. También me resulta como un resultado que se va un poco de tema, por eso creo que debería estar mejor contada esta parte para poder incluirla)}

@@

Figure~\ref{fig:sst-psi-2} shows regression maps between cEOF2 and Sea Surface Temperatures (SST) Streamfunction at 200 hPa. The Imaginary cEOF2 is associated with strong positive SST anomalies on the Central Pacific and negative anomalies on an area across the North of Australia and New Zealand, the South Pacific Convergence Zone (SPCZ) (Figure~\ref{fig:sst-psi-2}.a2). This pattern is almost canonically positive ENSO and indeed, the correlation between the Imaginary cEOF2 and the Oceanic Niño Index \citep{bamston1997} is very high 0.76 (CI: 0.59 -- 0.86). Streamfunction anomalies show a coherent picture. The Imaginary cEOF2 is associated with strong wave-like streamfunction anomalies emanating from the tropics (Figure~\ref{fig:sst-psi-2}.b2). This is consistent with what we know of the effect of ENSO on the extratropics: SST anomalies initiate anomalous convection that excites Rossby waves that propagate meridionally towards higher latitudes \textcolor{red}{(REFS??)}.

Since the Real cEOF2 represents just a different phase of the same wave train, one would expect that it would show a similar forcing pattern to the Imaginary cEOF with a slight translation of its location. However, Figure~\ref{fig:sst-psi-2}.a1 and b1 show that the Real cEOF2 is not associated either with any significant SST nor streamfunction anomalies in the tropics. The correlation between the Real cEOF2 and ENSO is also not significant (-0.0017 (CI: -0.31 -- 0.31)) \textcolor{red}{(No aparece el valor en el PDF)}. This lack of tropical signal suggests different natures of the different phases of the cEOF2 wave train.

\begin{figure}
\centering
\includegraphics{../figures/enso-phase-1.pdf}
\caption{\label{fig:enso-phase}Relationship between ENSO and phase of cEOF2 for the period 1979 -- 2019. Colours denote years with magnitude of cEOF2 greater or smaller than the 50th percentile. Black line is the fit ONI \textasciitilde{} cos(phase) + sin(phase) computed by OLS weighted by the magnitude of cEOF2.}
\end{figure}

To better explore the relationship between tropical forcing and phase of the cEOF2, Figure~\ref{fig:enso-phase} plots the ONI index and the phase of the cEOF2 for each SON trimester between 1979 and 2019, highlighting years in which the magnitude of cEOF2 is above the median. In years with positive ONI, the phase of the cEOF2 is mostly around +90º (corresponding with positive imaginary part) and vice versa. In years with near neutral ENSO, the phase of the cEOF2 is much more variable. The black line in Figure~\ref{fig:enso-phase} is an harmonical fit of the relationship between ONI and cEOF2 phase. The \(r^2\) corresponding to the fit is 0.56, with p-value \textless{} 0.001, indicating a quasi-sinusoidal relation between these two variables.

@@

Figure~\ref{fig:pp-america} shows regression maps of seasonal precipitation with each cEOF in South America. cEOF1 is associated with slight decreases in precipitation in Southern Brazil and Paraguay, although these are not statistically significant \ref{fig:pp-america} column a. The strongest precipitation anomalies are the ones associated with the Imaginary cEOF2. The positive anomalies on Southeastern South America (SESA) and Central Chile, and negative anomalies over South Atlantic Convergence Zone is a well known springtime precipitation signature of ENSO \citep{cai2020a} and it is also similar to the precipitation anomalies associated with the A-SAM \citep{campitelli2021}. This is not surprising considering the close relationship between the ONI, the Asymmetric SAM index and the Imaginary cEOF2 shown previously, but further consolidates the identification of this mode with the PSA pattern. The Real cEOF2, on the other hand, is associated with negative precipitation anomalies in a smaller area of SESA. There is a phase dependence of the precipitation anomalies in SESA (not shown), which resembles the relationship between ONI and the phase of cEOF2. \textcolor{red}{(Podrías agregar acá el coeficiente de determinación)}

@@

The precipitation anomalies in Oceania associated with cEOF1 are also weak and not statistically significant (Figure~\ref{fig:pp-oceania} column a) whereas EFO2 is associated with large and expansive anomalies. The Real cEOF2 is associated with positive seasonal anomalies in Northern and Eastern Australia, and negative ones over New Zealand, while precipitation anomalies associated with Imaginary cEOF2 are negative over all Eastern Australia. Again, these negative anomalies are similar to the springtime precipitation anomalies observed in relation with the Asymmetric SAM \citep{campitelli2021}.
\textcolor{red}{(Habría que agregar alguna referencia del ENSO que es más conocido)}

\begin{figure}
\centering
\includegraphics{../figures/australia-pp-phase-1.pdf}
\caption{\label{fig:australia-pp-phase}Same as Figure~\ref{fig:enso-phase} but for precipitacion anomalies averaged between 130°E and 155°E and 40°S and 15°S (box shown in Figure~\ref{fig:pp-oceania} column b).}
\end{figure}

To better understand the relationship between precipitation anomalies over Western Australia and the cEOF2, Figure~\ref{fig:australia-pp-phase} plots precipitation anomalies averaged over Eastern Australia (box in Figure~\ref{fig:pp-oceania}) as a function of cEOF2 phase. The black line is the equation \(\mathrm{Precipitation} = -0.26\sin{(\mathrm{phase})} + 0.22\cos{(\mathrm{phase})}\), whose coefficient we fitted by weighted least squares using the magnitude of the EFO2 as weight.
\textcolor{red}{(Como te puse antes la ecuación como mucho la pondría sobre la figura)}
The \(r^2\) corresponding to the fit is 0.37, with p-value \textless{} 0.001. There is a strong relationship between the phase of the cEOF2 and precipitation anomalies in Eastern Australia, but Real and Imaginary phases we chose are not exactly aligned with the direction that maximises this relationship. If one were to perform a more in-depth analysis of the relationship between this pattern and precipitation in this region, one would conceivably align one of the axis (either the Real or the Imaginary) to the one that maximises it.

>>>>

@@

Figure~\ref{fig:sst-psi-2} shows regression maps between cEOF2 and Sea Surface Temperatures (SST) Streamfunction at 200 hPa. The Imaginary cEOF2 is associated with strong positive SST anomalies on the Central Pacific and negative anomalies on an area across the North of Australia and New Zealand, the South Pacific Convergence Zone (SPCZ) (Figure~\ref{fig:sst-psi-2}.a2). This pattern is almost canonically positive ENSO and indeed, the correlation between the Imaginary cEOF2 and the Oceanic Niño Index \citep{bamston1997} is very high 0.76 (CI: 0.59 -- 0.86). Streamfunction anomalies show a coherent picture. The Imaginary cEOF2 is associated with strong wave-like streamfunction anomalies emanating from the tropics (Figure~\ref{fig:sst-psi-2}.b2). This is consistent with what we know of the effect of ENSO on the extratropics: SST anomalies initiate anomalous convection that excites Rossby waves that propagate meridionally towards higher latitudes \citep{mo2000}.

Since the Real cEOF2 represents just a different phase of the same wave train, one would expect that it would show a similar forcing pattern to the Imaginary cEOF with a slight translation of its location. However, Figure~\ref{fig:sst-psi-2}.a1 and b1 show that the Real cEOF2 is not associated either with any significant SST nor streamfunction anomalies in the tropics. The correlation between the Real cEOF2 and ENSO is also not significant (-0.0017 (CI: -0.31 -- 0.31)). This lack of tropical signal suggests different natures of the different phases of the cEOF2 wave train.

\begin{figure}
\centering
\includegraphics{../figures/enso-phase-1.pdf}
\caption{\label{fig:enso-phase}Relationship between ENSO and phase of cEOF2 for the period 1979 -- 2019. Colours denote years with magnitude of cEOF2 greater or smaller than the 50th percentile. Black line is the fit ONI \textasciitilde{} sin(phase) computed by OLS weighted by the magnitude of cEOF2.}
\end{figure}

To better explore the relationship between tropical forcing and phase of the cEOF2, Figure~\ref{fig:enso-phase} plots the ONI index and the phase of the cEOF2 for each SON trimester between 1979 and 2019, highlighting years in which the magnitude of cEOF2 is above the median. In years with positive ONI, the phase of the cEOF2 is mostly around +90º (corresponding with positive imaginary part) and vice versa. In years with near neutral ENSO, the phase of the cEOF2 is much more variable. The black line in Figure~\ref{fig:enso-phase} is a sinusoidal fit of the relationship between ONI and cEOF2 phase. The \(r^2\) corresponding to the fit is 0.56, with p-value \textless{} 0.001, indicating a quasi-sinusoidal relation between these two variables.

@@

Figure~\ref{fig:pp-america} shows regression maps of seasonal precipitation with each cEOF in South America. cEOF1 is associated with slight decreases in precipitation in Southern Brazil and Paraguay, although these are not statistically significant \ref{fig:pp-america} column a. The strongest precipitation anomalies are the ones associated with the Imaginary cEOF2. The positive anomalies on Southeastern South America (SESA) and Central Chile, and negative anomalies over South Atlantic Convergence Zone is a well known springtime precipitation signature of ENSO \citep{cai2020a} and it is also similar to the precipitation anomalies associated with the A-SAM \citep{campitelli2021}. This is not surprising considering the close relationship between the ONI, the Asymmetric SAM index and the Imaginary cEOF2 shown previously, but further consolidates the identification of this mode with the PSA pattern. The Real cEOF2, on the other hand, is associated with negative precipitation anomalies in a smaller area of SESA. There is a phase dependence of the precipitation anomalies in SESA (not shown), which resembles the relationship between ONI and the phase of cEOF2. A sinusoidal fit has an \(r^2\)~of~0.27 with p-value~= 0.002.

@@

The precipitation anomalies in Oceania associated with cEOF1 are also weak and not statistically significant (Figure~\ref{fig:pp-oceania} column a) whereas EFO2 is associated with large and expansive anomalies. The Real cEOF2 is associated with positive seasonal anomalies in Northern and Eastern Australia, and negative ones over New Zealand, while precipitation anomalies associated with Imaginary cEOF2 are negative over all Eastern Australia. Again, these negative anomalies are similar to the springtime precipitation anomalies observed in relation with the Asymmetric SAM \citep{campitelli2021} and ENSO \citep{cai2011}.

\begin{figure}
\centering
\includegraphics{../figures/australia-pp-phase-1.pdf}
\caption{\label{fig:australia-pp-phase}Same as Figure~\ref{fig:enso-phase} but for precipitation anomalies averaged between 130°E and 155°E and 40°S and 15°S (box shown in Figure~\ref{fig:pp-oceania} column b).}
\end{figure}

To better understand the relationship between precipitation anomalies over Western Australia and the cEOF2, Figure~\ref{fig:australia-pp-phase} plots precipitation anomalies averaged over Eastern Australia (box in Figure~\ref{fig:pp-oceania}) as a function of cEOF2 phase. The black line shows a sinusoidal fit, whose coefficients we fitted by weighted least squares using the magnitude of the EFO2 as weight. The \(r^2\) corresponding to the fit is 0.37, with p-value \textless{} 0.001. There is a strong relationship between the phase of the cEOF2 and precipitation anomalies in Eastern Australia, but Real and Imaginary phases we chose are not exactly aligned with the direction that maximises this relationship. If one were to perform a more in-depth analysis of the relationship between this pattern and precipitation in this region, one would conceivably align one of the axis (either the Real or the Imaginary) to the one that maximises it.
